Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-05-29 Manuscript.Rmd
modeling section done

  The resulting parameters were compared across cost groups (in Experiment 1) and across experiments using ANOVAs, and the $c_\text{demand}$ biases computed in Experiment 1 were compared against 0 using one-sample t-tests.

@@

# plot BIC across models for both exps
bicPlot <- ggplot(data = modelComparison, aes(Model, BIC, group = Exp, color = Exp)) +
  geom_line(size = 2) +
  geom_point(size = 4) +
  scale_color_grey(start = 0.2, end = 0.6, name = "", labels = c("Experiment 1", "Experiment 2")) +
  xlab("") +
  scale_x_discrete(labels = c("Base", "Adaptive", "Biased", "A + B")) +
  theme(legend.position = c(0.8, 0.8),
        #axis.text.x = element_text(angle = 20, hjust = 1),
        legend.key=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

@@

  The fourth model produced the lowest BIC in both experiments (Fig. 9A. Between-subjects = `r round(modelComparison[4, 3], digits = 2)`; Within-subjects = `r round(modelComparison[8, 3], digits = 2)`), indicating that the tenets of the Marginal Value Theorem were necessary but not sufficient to explain the observed adaptive behavior. Fig. 9B-C show parameter fits from the winning model for both experiments. The learning rate $\alpha$ and the bias $c_\text{demand}$ were of particular interest, as these parameters distinguished whether different demands globally affected the integration of the environmental richness or focally modulated an offer's perceived cost, respectively. Analyses of variance showed significant differences among Experiment 1 groups only for $c_\text{demand}$ (F(`r aovBtw$cost[[1]]$Df[1]`, `r aovBtw$cost[[1]]$Df[2]`) = `r fitFvals["cost"]`, *p* `r report_p(fitPvals["cost"])`), with $\alpha$ being uniformly low for all groups. Pairwise post-hoc permutations confirmed that the cognitive group (mean = `r round(mFits_btw["cost", "cogTask"], digits = 2)`, SD = `r round(sdFits_btw["cost", "cogTask"], digits = 2)`) had lower biases than either wait (mean = `r round(mFits_btw["cost", "wait"], digits = 2)`, SD = `r round(sdFits_btw["cost", "wait"], digits = 2)`; *p* `r report_p(fitPerms["Cognitive", "Wait"])`; Cohen's D = `r fitES["Cognitive", "Wait"]`) or physical (mean = `r round(mFits_btw["cost", "phys"], digits = 2)`, SD = `r round(sdFits_btw["cost", "phys"], digits = 2)`; *p* `r report_p(fitPerms["Cognitive", "Physical"])`; Cohen's D = `r fitES["Cognitive", "Physical"]`) groups. These biases were significantly higher than zero for the wait group (t(`r fitOStest$tresults[[1]]$parameter`) = `r round(fitOStest$tresults[[1]]$statistic, digits = 2)`, *p* `r report_p(fitOStest$tresults[[1]]$p.value)`), and lower than zero for the cognitive group (t(`r fitOStest$tresults[[2]]$parameter`) = `r round(fitOStest$tresults[[2]]$statistic, digits = 2)`, *p* `r report_p(fitOStest$tresults[[2]]$p.value)`). This suggests that rather than reducing costs, cognitive tasks boosted the value of their concomitant offers.

@@

# modelPlot <- (bicPlot + costPlot) / allFitsPlot
# ggsave("./Images/FIG9_modelresults.png", modelPlot, device = "png", height = 8, width = 11)
include_graphics("./Images/FIG9_modelresults.png")

@@

  The mean values of $c_\text{demand}$ from each group were carried over to the within-subjects fitting, and were aggregated over time according to a weighted decay into a final $c_\text{global}$ value. As can be seen on Fig. 9C, parameter values were comparable across experiments, pointing to consistency in the way decision makers approached our foraging scenarios. In addition, values of $\alpha_w$ suggested that $c_\text{demand}$ estimates converged towards the second half of the experiment (mean = `r mFits_wth[4]`, SD = `r sdFits_wth[4]`), gradually reducing bias by the end of the experimental session ($c_\text{global}$ mean = `r mFits_wth[3]`, SD = `r sdFits_wth[3]`). Model fits provided good qualitative resemblance to the observed choice dynamics in each experiment. This is illustrated in Fig. 9, in which 95% confidence intervals from within-subject fits are overlaid to empirical choices (Experiment 1 fits matched the left side of this figure, as expected).

>>>>

  We fit four free parameters to Experiment 1 data (the temperature $\beta$, learning rate $\alpha$, initial value of $\gamma$, and demand-specific bias $c_\text{demand}$) and five free parameters to Experiment 2 data (the temperature $\beta$, learning rate $\alpha$, initial value of $\gamma$, consensus bias $c_\text{global}$, and the rate of bias integration $\alpha_w$). The resulting parameters were compared across cost groups (in Experiment 1) and across experiments using ANOVAs, and the $c_\text{demand}$ biases computed in Experiment 1 were compared against 0 using one-sample t-tests.

@@

# plot BIC across models for both exps
bicPlot <- ggplot(data = modelComparison, aes(Model, BIC, group = Exp, color = Exp)) +
  geom_line(size = 2) +
  geom_point(size = 4) +
  scale_color_grey(start = 0.2, end = 0.6, name = "", labels = c("Experiment 1", "Experiment 2")) +
  xlab("") +
  ylab("iBIC") +
  scale_x_discrete(labels = c("Base", "Adaptive", "Biased", "A + B")) +
  theme(legend.position = c(0.8, 0.8),
        #axis.text.x = element_text(angle = 20, hjust = 1),
        legend.key=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

@@

  The fourth model produced the lowest iBIC in both experiments (Fig. 9A; Experiment 1 iBIC = `r round(modelComparison[4, 3], digits = 2)`; Experiment 2 iBIC = `r round(modelComparison[8, 3], digits = 2)`), indicating that the tenets of the Marginal Value Theorem were necessary but not sufficient to explain the observed adaptive behavior. Fig. 9B-C show parameter fits from the winning model for both experiments. The learning rate $\alpha$ and the bias $c_\text{demand}$ were of particular interest, as these parameters distinguished whether different forms of demand globally affected the integration of the environmental richness or directly modulated an offer's perceived cost, respectively. Analyses of variance showed significant differences among Experiment 1 groups only for $c_\text{demand}$ (F(`r aovBtw$cost[[1]]$Df[1]`, `r aovBtw$cost[[1]]$Df[2]`) = `r fitFvals["cost"]`, *p* `r report_p(fitPvals["cost"])`), with $\alpha$ being similar for all groups. Pairwise post-hoc permutations confirmed that the Cognitive group (mean = `r round(mFits_btw["cost", "cogTask"], digits = 2)`, SD = `r round(sdFits_btw["cost", "cogTask"], digits = 2)`) had lower values for the bias parameter than either the Wait group (mean = `r round(mFits_btw["cost", "wait"], digits = 2)`, SD = `r round(sdFits_btw["cost", "wait"], digits = 2)`; *p* `r report_p(fitPerms["Cognitive", "Wait"])`; Cohen's D = `r fitES["Cognitive", "Wait"]`) or the Physical group (mean = `r round(mFits_btw["cost", "phys"], digits = 2)`, SD = `r round(sdFits_btw["cost", "phys"], digits = 2)`; *p* `r report_p(fitPerms["Cognitive", "Physical"])`; Cohen's D = `r fitES["Cognitive", "Physical"]`). Parameter estimates were significantly greater than zero for the Wait group (t(`r fitOStest$tresults[[1]]$parameter`) = `r round(fitOStest$tresults[[1]]$statistic, digits = 2)`, *p* `r report_p(fitOStest$tresults[[1]]$p.value)`), and marginally significantly lower than zero for the Cognitive group (t(`r fitOStest$tresults[[2]]$parameter`) = `r round(fitOStest$tresults[[2]]$statistic, digits = 2)`, *p* `r report_p(fitOStest$tresults[[2]]$p.value)`).

@@

modelPlot <- (bicPlot + costPlot) / allFitsPlot
ggsave("./Images/FIG9_modelresult_part.png", modelPlot, device = "png", height = 8, width = 11)
include_graphics("./Images/FIG9_modelresults.png")

@@

  The mean values of $c_\text{demand}$ from each group were carried over to the model of Experiment 2, and were aggregated over time according to a weighted decay into a final $c_\text{global}$ value. As can be seen in Fig. 9C, other parameter values were comparable across experiments, pointing to consistency in the way decision makers approached our foraging scenarios. In addition, values of $\alpha_w$ suggested that $c_\text{demand}$ estimates converged towards the second half of the experiment (mean = `r mFits_wth[4]`, SD = `r sdFits_wth[4]`), gradually reducing bias by the end of the experimental session ($c_\text{global}$ mean = `r mFits_wth[3]`, SD = `r sdFits_wth[3]`). Model fits provided good qualitative resemblance to the observed choice dynamics in each experiment. This is illustrated in Fig. 9, in which 95% confidence intervals from within-subject fits are overlaid on empirical choices from Experiment 2.
